---
title: 'Assignment 8: Dynamic Game'
author: "Kohei Kawaguchi"
date: "2019/1/29"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include = FALSE}
library(EmpiricalIO)
library(magrittr)
library(stargazer)
library(knitr)
library(foreach)
library(ggplot2)
library(latex2exp)
library(codetools)
library(doParallel)
registerDoParallel()
knitr::opts_chunk$set(echo = TRUE, fig.align = "center")
```

## Simulate data

Suppose that there are $m = 1, \cdots, M$ markets and in each market there are $i = 1, \cdots, N$ firms and each firm makes decisions for $t = 1, \cdots, \infty$. In the following, I suppress the index of market, $m$. We solve the model under the infinite-horizon assumption, but generate data only for $t = 1, \cdots, T$. There are $L = 3$ state $\{1, 2, 3\}$ states for each each. Each firm can choose $K + 1 = 2$ actions $\{0, 1\}$. Thus, $m_a := (K + 1)^N$ and $m_s = L^N$. Let $a_i$ and $s_i$ be firm $i$'s action and state and $a$ and $s$ are vectors of individual actions and states. I write the code using matrix and loop so that we can translate easily to C++. Thus, the following code is not optimized for R.

The mean period payoff to firm $i$ is:
$$
\pi_i(a, s) := \tilde{\pi}(a_i, s_i, \overline{s}) :=  \alpha \ln s_i - \eta \ln \overline{s} - \beta a_i,
$$
where $\overline{s} := \frac{\sum_{i = 1}^N s_i}{N}$, $\alpha, \beta, \eta> 0$, and $\alpha > \eta$. The period payoff is:
$$
\tilde{\pi}(a_i, s_i, \overline{s})+ \epsilon_i(a_i),
$$
and $\epsilon_i(a_i)$ is an i.i.d. type-I extreme random variable that is independent of all the other variables.

At the beginning of each period, the state $s$ is realised and publicly observed. Then choice-specific shocks $\epsilon_i(a_i), a_i = 0, 1$ are realized and privately observed by firm $i = 1, \cdots, N$. Then each firm simultaneously chooses her action. Then, the game moves to next period. 

State transition is independent across firms conditional on individual state and action.

Suppose that $s_i > 1$ and $s_i < L$. If $a_i = 0$, the state stays at the same state with probability $1 - \kappa$ and moves down by 1 with probability $\kappa$. If $a = 1$, the state moves up by 1 with probability $\gamma$, moves down by 1 with probability $\kappa$, and stays at the same with probability $1 - \kappa - \gamma$. 

Suppose that $s_i = 1$. If $a_i = 0$, the state stays at the same state with probability 1. If $a_i = 1$, the state moves up by 1 with probability $\gamma$ and stays at the same with probability $1 - \gamma$. 

Suppose that $s_i = L$. If $a_i = 0$, the state stays at the same state with probability $1 - \kappa$ and moves down by 1 with probability $\kappa$. If $a = 1$, the state moves down by 1 with probability $\kappa$, and stays at the same with probability $1 - \kappa$. 



The mean period profit is summarized in $\Pi$ as:
$$
\Pi :=
\begin{pmatrix}
\pi(1, 1)\\
\vdots\\
\pi(m_a, 1)\\
\vdots \\
\pi(1, m_s)\\
\vdots\\
\pi(m_a, m_s)\\
\end{pmatrix}
$$

The transition law is summarized in $G$ as:

$$
g(a, s, s') := \mathbb{P}\{s_{t + 1} = s'|s_t = s, a_t = a\},
$$

$$
G := 
\begin{pmatrix}
g(1, 1, 1) & \cdots & g(1, 1, m_s)\\
\vdots & & \vdots \\
g(m_a, 1, 1) & \cdots & g(m_a, 1, m_s)\\
& \vdots & \\
g(1, m_s, 1) & \cdots & g(1, m_s, m_s)\\
\vdots & & \vdots \\
g(m_a, m_s, 1) & \cdots & g(m_a, m_s, m_s)\\
\end{pmatrix}.
$$
The discount factor is denoted by $\delta$. We simulate data for $M$ markets with $N$ firms for $T$ periods.

1. Set constants and parameters as follows:
```{r}
# set seed
set.seed(1)
# set constants 
L <- 3
K <- 1
T <- 100
N <- 3
M <- 100
lambda <- 1e-10
# dimension
m_a <- (K + 1)^N
m_s <- L^N
# set parameters
alpha <- 0.5
eta <- 0.1
beta <- 3
kappa <- 0.1
gamma <- 0.6
delta <- 0.95
```

2. Write a function `compute_action_state_space(K, L, N)` that returns a matrix for action and state space. Returned objects are list of matrix `A` and `S`, where `A` is a $m_a \times N$ matrix where $(k, i)$ element refers to the action taken by firm $i$ in $k$-th action profile and `S` is a `m_s \times N` matrix where $(l, i)$ refers to the state of firm $i$ in $l$-th state profile.

```{r}
output <- compute_action_state_space(K, L, N)
A <- output$A
head(A)
tail(A)
S <- output$S
head(S)
tail(S)
```


3. Write function `compute_PI_game(alpha, beta, eta, L, K, N)` that computes a list of $\Pi_i$ given parameters. Compute the list of true $\Pi$ under the true parameters. Don't use methods in `dplyr` and deal with matrix operations.

```{r}
PI <- compute_PI_game(alpha, beta, eta, L, K, N)
head(PI[[N]])
tail(PI[[N]])
```

4. Write function `compute_G_game(kappa, gamma, L, K, N)` that computes $G$ given parameters. Compute the true $G$ under the true parameters. Pick up several rows and manually compute the corresponding probabilities and test if the output of the function coincide. Don't use methods in `dplyr` and deal with matrix operations.

```{r}
G <- compute_G_game(kappa, gamma, L, K, N)
G[1, ]
G[m_a, ]
A[m_a,]
G[m_a, 1] - (1 - gamma)^N
G[m_a, 2] - gamma * (1 - gamma)^(N - 1)
G[m_a, L + 1] - gamma * (1 - gamma)^(N - 1)
G[m_a + 1, ]
A[1, ]
S[2, ]
S[1, ]
G[m_a + 1, 1] - kappa 
```

The exante-value function for a firm is written as a function of a conditional choice probability as follows:
$$
\varphi_i^{(\theta_1, \theta_2)}(p) := [I - \delta \Sigma_i(p) G]^{-1}[\Sigma_i(p)\Pi_i + D_i(p)],
$$
where $\theta_1 = (\alpha, \beta, \eta)$ and $\theta_2 = (\kappa, \gamma)$, $p_i(a_i|s)$ is the probability that firm $i$ choose action $a_i$ when the state profile is $s$, and:
$$
p(a|s) = \prod_{i = 1}^N p_i(a_i|s), 
$$
$$
p(s) = 
\begin{pmatrix}
p(1|s) \\
\vdots \\
p(m_a|s)
\end{pmatrix},
$$
$$
p = 
\begin{pmatrix}
p(1)\\
\vdots\\
p(m_s)
\end{pmatrix},
$$

$$
\Sigma(p) =
\begin{pmatrix}
p(1)' & & \\
 & \ddots & \\
 & & p(L)'
\end{pmatrix}
$$
and:
$$
\begin{equation}
D_i(p) =
\begin{pmatrix}
\sum_{k = 0}^K \mathbb{E}\{\epsilon_i^k|a_i = k\}p_i(a_i = k|1)\\
\vdots\\
\sum_{k = 0}^K \mathbb{E}\{\epsilon_i^k|a_i = k\}p_i(a_i = k|m_s)
\end{pmatrix}.
\end{equation}
$$

5. 

```{r}
p <- matrix(rep(1/(K + 1), N * m_a * m_s))
# compute marginal conitional choice probability
compute_marginal_p <-
  function(p, L, K, N) {
    p_marginal <-
      foreach (i = 1:N) %do% {
        
      }
    
  }


compute_D <-
  function(p, L, K, N) {
    # compute marginal conditional choice proabbility
    
    
  }
```


5. Write a function `compute_exante_value_game(p, PI, G, L, K, delta)` that returns the exante value function given a conditional choice probability. Don't use methods in `dplyr` and deal with matrix operations. When a choice probability is zero at some element, the corresponding element of $E(p)$ can be set at zero, because anyway we multiply the zero probability to the element and the corresponding element in $E(p)$ does not affect the result.

<!-- ```{r} -->
<!-- p <- matrix(rep(0.5, L * (K + 1)), ncol = 1); p -->
<!-- V <- compute_exante_value(p, PI, G, L, K, delta); V -->
<!-- ``` -->

<!-- The optimal conditional choice probability is written as a function of an exante value function as follows: -->
<!-- $$ -->
<!-- \Lambda^{(\theta_1, \theta_2)}(V)(a, s) := \frac{\pi(a, s) + \delta \sum_{s'}V(s')g(a, s, s')}{\sum_{a'}[\pi(a', s) + \delta \sum_{s'}V(s')g(a', s, s')]}, -->
<!-- $$ -->
<!-- where $V$ is an exante value function. -->

<!-- 4. Write a function `compute_ccp(V, PI, G, L, K, delta)` that returns the optimal conditional choice probability given an exante value function. Don't use methods in `dplyr` and deal with matrix operations. To do so, write a function `compute_choice_value(V, PI, G, delta)` that returns the choice-specific value function. Use this for debugging by checking if the results are intuitive. -->

<!-- ```{r} -->
<!-- value <- compute_choice_value(V, PI, G, delta); value -->
<!-- p <- compute_ccp(V, PI, G, L, K, delta); p -->
<!-- ``` -->

<!-- 5. Write a function that find the equilibrium conditional choice probability and ex-ante value function by iterating the update of an exante value function and an optimal conditional choice probability. The iteration should stop when $\max_s|V^{(r + 1)}(s) - V^{(r)}(s)| < \lambda$ with $\lambda = 10^{-10}$. -->

<!-- ```{r} -->
<!-- output <- solve_dynamic_decision(PI, G, L, K, delta, lambda); output -->
<!-- p <- output$p -->
<!-- V <- output$V -->
<!-- value <- compute_choice_value(V, PI, G, delta); value -->
<!-- ``` -->

<!-- 6. Write a function `simulate_dynamic_decision(p, s, PI, G, L, K, T, delta, seed)` that simulate the data for a single firm starting from an initial state for $T$ periods. The function should accept a value of seed and set the seed at the beginning of the procedure inside the function, because the process is stochastic. -->

<!-- ```{r} -->
<!-- # set initial value -->
<!-- s <- 1 -->
<!-- # draw simulation for a firm -->
<!-- seed <- 1 -->
<!-- df <- simulate_dynamic_decision(p, s, PI, G, L, K, T, delta, seed); df -->
<!-- ``` -->

<!-- 7. Write a function `simulate_dynamic_decision_across_firms(p, s, PI, G, L, K, T, N, delta)` that returns simulation data for $N$ firm. For firm $i$, set the seed at $i$ -->

<!-- ```{r, eval = FALSE} -->
<!-- df <- simulate_dynamic_decision_across_firms(p, s, PI, G, L, K, T, N, delta) -->
<!-- save(df, file = "data/A7_df.RData") -->
<!-- ``` -->
<!-- ```{r} -->
<!-- load(file = "data/A7_df.RData") -->
<!-- df -->
<!-- ``` -->


<!-- 8. Write a function `estimate_ccp(df)` that returns a non-parametric estimate of the conditional choice probability in the data. Compare the estimated conditional choice probability and the true conditional choice probability by a bar plot. -->

<!-- ```{r} -->
<!-- p_est <- estimate_ccp(df) -->
<!-- check_ccp <- cbind(p, p_est) -->
<!-- colnames(check_ccp) <- c("true", "estimate") -->
<!-- check_ccp <- check_ccp %>% -->
<!--   reshape2::melt() -->
<!-- ggplot(data = check_ccp, aes(x = Var1, y = value,  -->
<!--                              fill = Var2)) + -->
<!--   geom_bar(stat = "identity", position = "dodge") + -->
<!--   labs(fill = "Value") + xlab("action/state") + ylab("probability") -->
<!-- ``` -->

<!-- 9. Write a function `estimate_G(df)` that returns a non-parametric estiamte of the transition matrix in the data. Compare the estimated transition matrix and the true transition matrix by a bar plot. -->


<!-- ```{r} -->
<!-- G_est <- estimate_G(df); G_est -->
<!-- check_G <- data.frame(type = "true", reshape2::melt(G)) -->
<!-- check_G_est <- data.frame(type = "estimate", reshape2::melt(G_est)) -->
<!-- check_G <- rbind(check_G, check_G_est) -->
<!-- check_G$variable = paste(check_G$Var1, check_G$Var2, sep = "_") -->
<!-- ggplot(data = check_G, aes(x = variable, y = value, -->
<!--                            fill = type)) + -->
<!--     geom_bar(stat = "identity", position = "dodge") + -->
<!--   labs(fill = "Value") + xlab("action/state/state") + ylab("probability") + -->
<!--   theme(axis.text.x = element_blank()) -->
<!-- ``` -->


<!-- ## Estimate parameters -->

<!-- 1. Vectorize the parameters as follows: -->

<!-- ```{r} -->
<!-- theta_1 <- c(alpha, beta) -->
<!-- theta_2 <- c(kappa, gamma) -->
<!-- theta <- c(theta_1, theta_2) -->
<!-- ``` -->


<!-- First, we estimate the parameters by a nested fixed-point algorithm. The loglikelihood for $\{a_{it}, s_{it}\}_{i = 1, \cdots, N, t = 1, \cdots, T}$ is: -->
<!-- $$ -->
<!-- \frac{1}{NT} \sum_{i = 1}^N \sum_{t = 1}^T[\log\mathbb{P}\{a_{it}|s_{it}\} + \log \mathbb{P}\{s_{i, t + 1}|a_{it}, s_{it}\}], -->
<!-- $$ -->
<!-- with $\mathbb{P}\{s_{i, T + 1}|a_{iT}, s_{iT}\} = 1$ for all $i$ as $s_{i, T + 1}$ is not observed.  -->

<!-- 2. Write a function `compute_loglikelihood_NFP(theta, df, delta, L, K)` that compute the loglikelihood. -->


<!-- ```{r} -->
<!-- loglikelihood <- compute_loglikelihood_NFP(theta, df, delta, L, K); loglikelihood -->
<!-- ``` -->

<!-- 3. Check the value of the objective function around the true parameter. -->

<!-- ```{r, eval = FALSE} -->
<!-- # label -->
<!-- label <- c("\\alpha", "\\beta", "\\kappa", "\\gamma") -->
<!-- label <- paste("$", label, "$", sep = "") -->
<!-- # compute the graph -->
<!-- graph <- foreach (i = 1:length(theta)) %do% { -->
<!--   theta_i <- theta[i] -->
<!--   theta_i_list <- theta_i * seq(0.8, 1.2, by = 0.05) -->
<!--   objective_i <-  -->
<!--     foreach (j = 1:length(theta_i_list), -->
<!--              .combine = "rbind") %do% { -->
<!--                theta_ij <- theta_i_list[j] -->
<!--                theta_j <- theta -->
<!--                theta_j[i] <- theta_ij -->
<!--                objective_ij <-  -->
<!--                  compute_loglikelihood_NFP( -->
<!--                    theta_j, df, delta, L, K); loglikelihood -->

<!--                return(objective_ij) -->
<!--              } -->
<!--   df_graph <- data.frame(x = theta_i_list, y = objective_i)  -->
<!--   g <- ggplot(data = df_graph, aes(x = x, y = y)) +  -->
<!--     geom_point() + -->
<!--     geom_vline(xintercept = theta_i, linetype = "dotted") + -->
<!--     ylab("objective function") + xlab(TeX(label[i])) -->
<!--   return(g) -->
<!-- } -->
<!-- save(graph, file = "data/A7_NFP_graph.RData") -->
<!-- ``` -->
<!-- ```{r} -->
<!-- load(file = "data/A7_NFP_graph.RData") -->
<!-- graph -->
<!-- ``` -->

<!-- 4. Estiamte the parameters by maximizing the loglikelihood. To keep the model to be well-defined, impose an ad hoc lower and upper bounds such that $\alpha \in [0, 1], \beta \in [0, 5], \kappa \in [0, 0.2], \gamma \in [0, 0.7]$. -->

<!-- ```{r, eval = FALSE} -->
<!-- lower <- rep(0, length(theta)) -->
<!-- upper <- c(1, 5, 0.2, 0.7) -->
<!-- NFP_result <- -->
<!--   optim(par = theta, -->
<!--         fn = compute_loglikelihood_NFP, -->
<!--         method = "L-BFGS-B", -->
<!--         lower = lower, -->
<!--         upper = upper, -->
<!--         control = list(fnscale = -1), -->
<!--         df = df,  -->
<!--         delta = delta, -->
<!--         L = L, -->
<!--         K = K) -->
<!-- save(NFP_result, file = "data/A7_NFP_result.RData") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- load(file = "data/A7_NFP_result.RData") -->
<!-- NFP_result -->
<!-- compare <- -->
<!--   data.frame( -->
<!--     true = theta, -->
<!--     estimate = NFP_result$par -->
<!--   ); compare -->
<!-- ``` -->

<!-- Next, we estimate the parameters by CCP approach. -->

<!-- 5. Write a function `estimate_theta_2(df)` that returns the estimates of $\kappa$ and $\gamma$ directly from data by counting relevant events. -->
<!-- ```{r} -->
<!-- theta_2_est <- estimate_theta_2(df); theta_2_est -->
<!-- ``` -->

<!-- The objective function of the minimum distance estimator based on the conditional choice probability approach is: -->
<!-- $$ -->
<!-- \frac{1}{KL}\sum_{s = 1}^L \sum_{a = 1}^K\{\hat{p}(a, s) - p^{(\theta_1, \theta_2)}(a, s)\}^2, -->
<!-- $$ -->
<!-- where $\hat{p}$ is the non-parametric estimate of the conditional choice probability and $p^{(\theta_1, \theta_2)}$ is the optimal conditional choice probability under parameters $\theta_1$ and $\theta_2$. -->

<!-- 6. Write a function `compute_CCP_objective(theta_1, theta_2, p_est, L, K, delta)` that returns the objective function of the above minimum distance estimator given a non-parametric estimate of the conditional choice probability and $\theta_1$ and $\theta_2$. -->


<!-- ```{r} -->
<!-- compute_CCP_objective(theta_1, theta_2, p_est, L, K, delta) -->
<!-- ``` -->

<!-- 3. Check the value of the objective function around the true parameter. -->

<!-- ```{r, eval = FALSE} -->
<!-- # label -->
<!-- label <- c("\\alpha", "\\beta") -->
<!-- label <- paste("$", label, "$", sep = "") -->
<!-- # compute the graph -->
<!-- graph <- foreach (i = 1:length(theta_1)) %do% { -->
<!--   theta_i <- theta_1[i] -->
<!--   theta_i_list <- theta_i * seq(0.8, 1.2, by = 0.05) -->
<!--   objective_i <-  -->
<!--     foreach (j = 1:length(theta_i_list), -->
<!--              .combine = "rbind") %do% { -->
<!--                theta_ij <- theta_i_list[j] -->
<!--                theta_j <- theta_1 -->
<!--                theta_j[i] <- theta_ij -->
<!--                objective_ij <-  -->
<!--                  compute_CCP_objective(theta_j, theta_2, p_est, L, K, delta) -->
<!--                return(objective_ij) -->
<!--              } -->
<!--   df_graph <- data.frame(x = theta_i_list, y = objective_i)  -->
<!--   g <- ggplot(data = df_graph, aes(x = x, y = y)) +  -->
<!--     geom_point() + -->
<!--     geom_vline(xintercept = theta_i, linetype = "dotted") + -->
<!--     ylab("objective function") + xlab(TeX(label[i])) -->
<!--   return(g) -->
<!-- } -->
<!-- save(graph, file = "data/A7_CCP_graph.RData") -->
<!-- ``` -->
<!-- ```{r} -->
<!-- load(file = "data/A7_CCP_graph.RData") -->
<!-- graph -->
<!-- ``` -->

<!-- 4. Estiamte the parameters by minimizing the objective function. To keep the model to be well-defined, impose an ad hoc lower and upper bounds such that $\alpha \in [0, 1], \beta \in [0, 5]$. -->

<!-- ```{r, eval = FALSE} -->
<!-- lower <- rep(0, length(theta_1)) -->
<!-- upper <- c(1, 5) -->
<!-- CCP_result <- -->
<!--   optim(par = theta_1, -->
<!--         fn = compute_CCP_objective, -->
<!--         method = "L-BFGS-B", -->
<!--         lower = lower, -->
<!--         upper = upper, -->
<!--         theta_2 = theta_2_est,  -->
<!--         p_est = p_est,  -->
<!--         L = L,  -->
<!--         K = K,  -->
<!--         delta = delta) -->
<!-- save(CCP_result, file = "data/A7_CCP_result.RData") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- load(file = "data/A7_CCP_result.RData") -->
<!-- CCP_result -->
<!-- compare <- -->
<!--   data.frame( -->
<!--     true = theta_1, -->
<!--     estimate = CCP_result$par -->
<!--   ); compare -->
<!-- ``` -->





