# Assignments {#assignments}

## Assignment 1

1. Setup R, RStudio, and LaTeX environment in your local computer.
1. 
    

## Assignment 2

### Sample Code

\paragraph{OLS, Panel, and IV Regressions} Read the data. The path has to be correctly specified relative to your working directotry.
```{r}
library(magrittr)
data <- readr::read_csv(file = "data/data_ps2.csv")
summary(data)
```
The variable `firm` is the firm identifier, and `year` is the period identifier. `Y` is the output of the period, `K` is the capital stock at the beginning of the priod, `L` is the number of labor employed during the period, and `I` is the investment of the period. If a firm exits from the market in a period, it does not show up in the data in the period.

Suppose that the production function of the firms takes the form of:
$$
Y_{jt} = A_{jt} L_{jt}^{\beta_L} K_{jt}^{\beta_K},
$$
where $A_{jt}$ is the unobserved productivity of firm $j$ in period $t$. We estimate $\beta_L$, $\beta_K$, and the distribution of $A_{jt}$ using the data above.

You first run OLS regression:
```{r}
ols <- lm(data = data, formula = log(Y) ~ log(L) + log(K))
```

Second, you run a panel regression with firm-specific unobserved fixed effects:
```{r}
panel <- plm::plm(data = data, formula = log(Y) ~ log(L) + log(K),
                  effect = "individual", model = "within", index = c("firm", "year"))
```
where I used a package `plm` for the panel regression. In this code, I wrote like `plm::plm` to call function `plm` in package `plm`. This works almost in the same way as first calling the library by `library(plm)` and then using the function `plm`. I prefer this way of loading library and calling function, because i) it is more explicit, ii) it avoids the conflicts in the namespace, and iii) it can be directly used in the multi-node parallel computing.

Third, you run an instrumental vavriable estimation using the lagged labor and capital inputs as the instruments for the labor and capital inputs:
```{r}
data <- data %>%
  dplyr::arrange(firm, year) %>%
  dplyr::group_by(firm) %>%
  dplyr::mutate(L1_L = dplyr::lag(L), L1_K = dplyr::lag(K),
                L2_L = dplyr::lag(L, 2), L2_K = dplyr::lag(K, 2)) %>%
  dplyr::ungroup()
lagged <- AER::ivreg(data = data, formula = log(Y) ~ log(L) + log(K) | 
                       log(L1_L) + log(L1_K) + log(L2_L) + log(L2_K))
```

\paragraph{Olley-Pakes Approach} Next, we estiamte the parameters using Olley-Pakes approach. The code for the first step is:
```{r, results = "hide"}
### compute the bandwidth for the nonparametric term 
### this takes a few minutes depending on your computer
bw <- np::npplregbw(formula = log(Y) ~ log(L) | log(K) + I, data = data)
### estimate the partially linear model with the bandwidth above
step1 <- np::npplreg(bws = bw)
```

Then, prepare some variables for the second step estimation:
```{r}
### extract the estimate of the log(L) coefficient
beta_l_hat <- coef(step1)
beta_l_se <- sqrt(vcov(step1))
### compute relevant values
data <- data %>%
  dplyr::mutate(phi = predict(step1) - beta_l_hat*log(L)) %>%
  dplyr::arrange(firm, year) %>%
  dplyr::group_by(firm) %>%
  dplyr::mutate(L1_phi = lag(phi)) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(ly = log(Y) - beta_l_hat*log(L))
### extract the estimate of the log(K) coefficient from panel regression
beta_panel <- coef(panel)
```

Finally, you can run the second step of Olley-Pakes approach:
```{r}
## second step
step2 <- nls(formula = ly ~ beta_0 + beta_k*log(K) + 
              b_1*(L1_phi - beta_0 - beta_k*log(L1_K)) +
              b_2*(L1_phi - beta_0 - beta_k*log(L1_K))^2 +
              b_3*(L1_phi - beta_0 - beta_k*log(L1_K))^3 +
              b_4*(L1_phi - beta_0 - beta_k*log(L1_K))^4,
              data = data,
              start = list(b_1 = 0, b_2 = 0, b_3 = 0, b_4 = 0,
                           beta_0 = 0, beta_k = beta_panel[2]))
```

The package `np` contains several functions for nonparametric and semiparametric estimations. The function `npplreg` is used for estimating a partially linear model. To do so, you first obtain a data-driven bandwidth by using the function `npplregbw`. The function `nls` is a basic function to estimate a nonlinear model. In this implementation, I am not taking into account the nonparametric issues in the second step for a pedagogical reason: the standard errors may be wrong but you do not have to compute the correct standard errors for the Olley-Pakes estimators in this problem set.


### Problems

All the results and the discussion should be summarized in a pdf to submit. The code can be either submitted independently of the pdf file as a r file or included in the pdf file using rmarkdown.

1. Replicate the codes above.
1. Summarise the estimation results in a table.
1. Discuss the difference in the estimates across estimators.
1. The Olley-Pakes estimator above does not control for the selection bias due to exit. Implement a Olley-Pakes estimator controlling for the selection bias.
1. Summarise the estimation result of the Olley-Pakes estimator controlling for the selection bias.
1. Discuss the difference in the estimates from the other estimates.


## Assignment 4

## Assignment 5

## Assignment 6

## Assignment 7

## Assignment 8
