# Assignment 1: Basic Programming in R {#assignment1}

```{r, include=FALSE}
library(EmpiricalIO)
library(magrittr)
library(foreach)
library(ggplot2)
library(latex2exp)
knitr::opts_chunk$set(echo = FALSE, fig.align = "center")
```

Report the following results in html format using R markdown. In other words, replicate this document. You write functions in a separate R file and put in `R` folder in the project folder. Build the project as a package and load it from the R markdown file. The execution code sholuld be written in R markdown file. 

You submit:

- R file containing functions.
- R markdown file containing your answers and executing codes.
- HTML(or PDF) report generated from the R markdown.

## Simulate data

Consider to simulate data from the following model and estimate the parameters from the simulated data.

$$
y_{ij} = 1\{j = \text{argmax}_{k = 1, 2} \beta x_k + \epsilon_{ik} \},
$$
where $\epsilon_{ik}$ follows i.i.d. type-I extreme value distribution, $\beta = 0.2$,  $x_1 = 0$ and $x_2 = 1$.

1. To simulate data, first make a data frame as follows:
```{r}
df <- 
  expand.grid(
    i = 1:1000, 
    k = 1:2
    ) %>%
  tibble::as_tibble() %>%
  dplyr::mutate(x = ifelse(k == 1, 0, 1)) %>%
  dplyr::arrange(
    i, 
    k
    )
```
```{r, echo = TRUE}
df
```
2. Second, draw type-I extreme value random variables. Set the seed at 1. You can use `evd` package to draw the variables. You should get exactly the same realization if the seed is correctly set.
```{r}
set.seed(1)
```

```{r}
df <- 
  df %>%
  dplyr::mutate(e = evd::rgev(dim(df)[1]))
```
```{r, echo = TRUE}
df
```
3. Third, compute the latent value of each option to obtain the following data frame:
```{r, echo = TRUE}
beta <- 0.2
```

```{r}
df <- 
  df %>%
  dplyr::mutate(latent = beta * x + e)
```
```{r, echo = TRUE}
df
```
4. Finally, compute $y$ by comparing the latent values of $k = 1, 2$ for each $i$ to obtain the following result:
```{r}
df <- 
  df %>%
  dplyr::group_by(i) %>%
  dplyr::mutate(y = ifelse(latent == max(latent), 1, 0)) %>%
  dplyr::ungroup()
```
```{r, echo = TRUE}
df
```

## Estimate the parameter

Now you generated simulated data. Suppose you observe $x_k$ and $y_{ik}$ for each $i$ and $k$ and estimate $\beta$ by a maximum likelihood estimator. The likelihood for $i$ to choose $k$ ($y_{ik} = 1$) can be shown to be:
$$
p_{ik}(\beta) = \frac{\exp(\beta x_k)}{\exp(\beta x_1) + \exp(\beta x_2)}.
$$

Then, the likelihood of observing $\{y_{ik}\}_{i, k}$ is:
$$
L(\beta) = \prod_{i = 1}^{1000} p_{i1}(\beta)^{y_{i1}} [1 - p_{i1}(\beta)]^{1 - y_{i1}},
$$
and the log likelihood is:
$$
l(\beta) = \sum_{i = 1}^{1000}\{y_{i1}\log p_{i1}(\beta) + (1 - y_{i1})\log [1 - p_{i1}(\beta)\}.
$$

1. Write a function to compute the livelihood for a given $\beta$ and data and name the function `loglikelihood_a1(b, df)`. 
```{r, echo = TRUE}
loglikelihood_a1(
    b = 1,
    df = df
    ) 
```

1. Compute the value of log likelihood for $\beta = 0, 0.1, \cdots, 1$ and plot the result using `ggplot2` packages. You can use `latex2exp` package to use LaTeX math symbol in the label:
```{r, echo = TRUE}
b_seq <- seq(0, 1, 0.1)
output <- 
  foreach (
    b = b_seq,
    .combine = "rbind"
    ) %do% {
      l <- 
        loglikelihood_a1(
          b, 
          df
          )
      return(l)
}
output <- 
  data.frame(
    x = b_seq, 
    y = output
    )
output %>%
  ggplot(
    aes(
      x = x, 
      y = y
      )
    ) +
  geom_point() + 
  xlab(TeX("$\\beta$")) + 
  ylab("Loglikelihood") +
  theme_classic()
```


1. Find and report $\beta$ that maximizes the log likelihood for the simulated data. You can use `optim` function to achieve this. You will use `Brent` method and set the lower bound at -1 and upper bound at 1 for the parameter search.
```{r}
result <- 
  optim(
    par = beta,
    fn = loglikelihood_a1,
    df = df,
    method = "Brent",
    lower = -1,
    upper = 1,
    control = list(fnscale = - 1)
    )
result
```









