# Assignment 9: Auction {#assignment9}

```{r, include = FALSE}
library(EmpiricalIO)
library(magrittr)
library(stargazer)
library(knitr)
library(foreach)
library(ggplot2)
library(latex2exp)
library(codetools)
library(doParallel)
registerDoParallel()
knitr::opts_chunk$set(cache = FALSE, echo = TRUE, fig.align = "center")
```

## Simulate data

We simulate bid data from a second- and first-price sealed bid auctions.

First, we draw bid data from a second-price sealed bid auctions. Suppose that for each auction $t = 1, \cdots, T$, there are $i = 2, \cdots, n_t$ potential bidders. At each auction, an auctioneer allocates one item and sets the reserve price at $r_t$. When the signal for bidder $i$ in auction $t$ is $x_{it}$, her expected value of the item is $x_{it}$. A signal $x_{it}$ is drawn from an i.i.d. beta distribution $B(\alpha, \beta)$. Let $F_X(\cdot; \alpha, \beta)$ be its distribution and $f_X(\cdot; \alpha, \beta)$ be the density. A reserve is set at 0.2. $n_t$ is drawn from a Poisson distribution with mean $\lambda$. If $n_t = 1$, replace with $n_t = 2$ to ensure at least two potential bidders.  An equilibrium strategy is such that a bidder participates and bids $\beta(x) = x$ if $x \ge r_t$ and does not participate otherwise.

1. Set the constants and parameters as follows:
```{r}
# set seed
set.seed(1)
# number of auctions
T <- 100
# parameter of value distribution
alpha <- 2
beta <- 2
# prameters of number of potential bidders
lambda <- 10
```

2. Draw a vector of valuations and reservation prices.

```{r, echo = TRUE}
# number of bidders
N <- 
  rpois(
    T, 
    lambda
    )
N <- 
  ifelse(
    N == 1, 
    2, 
    N
    )
# draw valuations
valuation <-
  foreach (
    tt = 1:T, 
    .combine = "rbind"
    ) %do% {
    n_t <- N[tt]
    header <- 
      expand.grid(
        t = tt,
        i = 1:n_t
        ) 
    return(header)
  }
valuation <- 
  valuation %>%
  tibble::as_tibble() %>%
  dplyr::mutate(x = 
                  rbeta(
                    length(i),
                    alpha,beta
                    )
                )
ggplot(
  valuation, 
  aes(x = x)
  ) + 
  geom_histogram(
    fill = "steelblue",
    alpha = 0.8
    )
# draw reserve prices
reserve <- 0.2
reserve <- 
  tibble::tibble(
    t = 1:T, 
    r = reserve
    )
```


3. Write a function `compute_winning_bids_second(valuation, reserve)` that returns a winning bid from each second-price auction. It returns nothing for an auction in which no bid was above the reserve price. In the output, `t` refers to the auction index, `m` to the number of actual bidders, `r` to the reserve price, and `w` to the winning bid.
```{r}
# compute winning bids from second-price auction
df_second_w <-
  compute_winning_bids_second(
    valuation = valuation, 
    reserve = reserve
    )
df_second_w
ggplot(
  df_second_w, 
  aes(x = w)
  ) + 
  geom_histogram(
    fill = "steelblue",
    alpha = 0.8
    )
```

Next, we simulate bid data from first-price sealed bid auctions. The setting is the same as the second-price auctions expect for the auction rule. An equilibrium bidding strategy is to participate and bid:
$$
\beta(x) = x - \frac{\int_{r_t}^x F_X(t)^{N - 1}}{F_X(x)^{N - 1}},
$$
if $x \ge r$ and not to participate otherwise.

4. Write a function `bid_first(x, r, alpha, beta, n)` that returns the equilibrium bid. To integrate a function, use `integrate` function in R. It returns 0 if $x < r$.
```{r}
# compute bid from first-price auction
n <- N[1]
m <- N[1]
x <- valuation[1, "x"] %>% 
  as.numeric();
x
r <- 
  reserve[1, "r"] %>% 
  as.numeric(); 
r
b <- 
  bid_first(
    x = x,
    r = r,
    alpha = alpha,
    beta = beta,
    n = n
    );
b
x <- r/2;
x
b <- 
  bid_first(
    x = x,
    r = r, 
    alpha = alpha,
    beta = beta, 
    n = n
    ); 
b
b <- 
  bid_first(
    x = 1, 
    r = r, 
    alpha = alpha, 
    beta = beta, 
    n = n
    ); 
b
```

5. Write a function `compute_bids_first(valuation, reserve, alpha, beta)` that returns bids from each first-price auctions. It returns bids below the reserve price.

```{r}
# compute bid data from first-price auctions
df_first <- 
  compute_bids_first(
    valuation = valuation, 
    reserve = reserve,
    alpha = alpha, 
    beta = beta
    )
df_first
ggplot(
  df_first,
  aes(x = b)
  ) + 
  geom_histogram(
    fill = "steelblue",
    alpha = 0.8
    )
```

6. Write a function `compute_winning_bids_first(valuation, reserve, alpha, beta)` that returns only the winning bids from each first-price auction. It will call `compute_bids_first` inside the function. It does not return anything when no bidder bids above the reserve price.

```{r}
# compute winning bids from first-price auctions
df_first_w <-
  compute_winning_bids_first(
    valuation = valuation,
    reserve = reserve,
    alpha = alpha,
    beta = beta
    )
df_first_w
ggplot(
  df_first_w, 
  aes(x = w)
  ) + 
  geom_histogram(
    fill = "steelblue", 
    alpha = 0.8
    )
```

## Estimate the parameters

We first estimate the parameters from the winning bids data of second-price auctions. We estimate the parameters by maximizing a log-likelihood.

$$
l(\alpha, \beta) := \frac{1}{T} \sum_{t = 1}^T \ln\frac{h_t(w_t)^{1\{m_t > 1\}} \mathbb{P}\{m_t = 1\}^{1\{m_t = 1\}}}{1 - \mathbb{P}\{m_t = 0\}},
$$
where:
$$
\mathbb{P}\{m_t = 0\} := F_X(r_t)^{n_t},
$$
$$
\mathbb{P}\{m_t = 1\} := n_t F_X(r_t; \alpha, \beta)^{n_t - 1} [1 - F_X(r_t; \alpha, \beta)].
$$

$$
h_t(w_t) := n_t (n_t - 1) F_X(w_t; \alpha, \beta)^{n_t - 2} [1 - F_X(w_t; \alpha, \beta)] f_X(w_t; \alpha, \beta).
$$

1. Write a function `compute_p_second_w(w, r, m, n, alpha, beta)` that computes $\mathbb{P}\{m_t = 1\}$ if $m_t = 1$ and $h_t(w_t)$ if $m_t > 1$.

```{r}
# compute probability density for winning bids from a second-price auction
w <- df_second_w[1, ]$w
r <- df_second_w[1, ]$r
m <- df_second_w[1, ]$m
n <- df_second_w[1, ]$n
compute_p_second_w(
  w = w, 
  r = r, 
  m = m,
  n = n,
  alpha = alpha,
  beta = beta
  )
```

2. Write a function `compute_m0(r, n, alpha, beta)` that computes $\mathbb{P}\{m_t = 0\}$. 

```{r}
# compute non-participation probability
compute_m0(
  r = r, 
  n = n, 
  alpha = alpha, 
  beta = beta
  )
```


2. Write a function `compute_loglikelihood_second_price_w(theta, df_second_w)` that computes the log-likelihood for a second-price auction winning bid data.

```{r}
# compute log-likelihood for winning bids from second-price auctions
theta <- 
  c(
    alpha, 
    beta
    )
compute_loglikelihood_second_price_w(
  theta = theta, 
  df_second_w = df_second_w
  )
```

3. Compare the value of objective function around the true parameters.

```{r}
# label
label <- 
  c(
    "\\alpha",
    "\\beta"
    )
label <- 
  paste(
    "$", 
    label, 
    "$",
    sep = ""
    )
# compute the graph
graph <- 
  foreach (
    i = 1:length(theta)
    ) %do% {
  theta_i <- theta[i]
  theta_i_list <- 
    theta_i * seq(
      0.8,
      1.2, 
      by = 0.05
      )
  objective_i <-
    foreach (
      j = 1:length(theta_i_list),
      .packages = c(
                    "EmpiricalIO",
                    "foreach",
                    "magrittr"
                   ),
      .combine = "rbind"
      ) %dopar% {
               theta_ij <- theta_i_list[j]
               theta_j <- theta
               theta_j[i] <- theta_ij
               objective_ij <- 
                 compute_loglikelihood_second_price_w(
                   theta_j, 
                   df_second_w
                   )
               return(objective_ij)
             }
  df_graph <- 
    data.frame(
      x = as.numeric(theta_i_list), 
      y = as.numeric(objective_i)
      )
  g <- 
    ggplot(
      data = df_graph, 
      aes(
        x = x,
        y = y
        )
      ) +
    geom_point() +
    geom_vline(
      xintercept = theta_i,
      linetype = "dotted"
      ) +
    ylab("objective function") + 
    xlab(TeX(label[i]))
  return(g)
}
saveRDS(
  graph, 
  file = "data/a9/A9_second_parametric_graph.rds"
  )
```
```{r}
graph <- 
  readRDS(file = "data/a9/A9_second_parametric_graph.rds")
graph
```

4. Estimate the parameters by maximizing the log-likelihood.

```{r}
result_second_parametric <-
  optim(
    par = theta,
    fn = compute_loglikelihood_second_price_w,
    df_second_w = df_second_w,
    method = "L-BFGS-B",
    control = list(fnscale = -1)
  )
saveRDS(
  result_second_parametric, 
  file = "data/A9_result_second_parametric.rds"
  )
```
```{r}
result_second_parametric <- 
  readRDS(file = "data/A9_result_second_parametric.rds")
result_second_parametric
comparison <-
  data.frame(
    true = theta,
    estimate = result_second_parametric$par
  )
comparison
```

Next, we estimate the parameters from the winning bids data from first-price auctions. We estimate the parameters by maximizing a log-likelihood.

5. Write a function `inverse_bid_equation(x, b, r, alpha, beta, n)` that returns $\beta(x) - b$ for a bid $b$. Write a function `inverse_bid_first(b, r, alpha, beta, n)` that is an inverse function `bid_first` with respect to the signal, that is,
$$
\eta(b) := \beta^{-1}(b).
$$

To do so, we can use a built-in function called `uniroot`, which solves $x$ such that $f(x) = 0$ for scalar $x$. In `uniroot`, `lower` and `upper` are set at $r_t$ and $\beta(1)$, respectively.

```{r}
r <- df_first_w[1, "r"] %>%
  as.numeric()
n <- df_first_w[1, "n"] %>%
  as.integer()
b <- 0.5 * r + 0.5 
x <- 0.5
# compute invecrse bid equation
inverse_bid_equation(
  x = x,
  b = b,
  r = r,
  alpha = alpha, 
  beta = beta,
  n = n
  )
# compute inverse bid
inverse_bid_first(
  b = b, 
  r = r,
  alpha = alpha,
  beta = beta,
  n = n
  )
```

The log-likelihood conditional on $m_t \ge 1$ is:
$$
l(\alpha, \beta) := \frac{1}{T}\sum_{t = 1}^T \log \frac{h_t(w_t)}{1 - F_X(r_t)^{n_t}},
$$
where the probability density of having $w_t$ is:
$$
\begin{split}
h_t(w_t) &= n_t F_X[\eta_t(w_t)]^{n_t - 1} f_X[\eta_t(w_t)] \eta_t'(w_t)\\
&= \frac{n_t F_X[\eta_t(w_t)]^{n_t}}{(n_t - 1)[\eta_t(w_t) - w_t]},
\end{split}
$$
where the second equation is from the first-order condition.


6. Write a function `compute_p_first_w(w, r, alpha, beta, n)` that returns $h_t(w)$. Remark that the equilibrium bid at specific parameters is `bid_first(1, r, alpha, beta, n)`. If the observed wining bid `w` is above the upper limit, the function will issue an error. Therefore, inside the function `compute_p_first_w(w, r, alpha, beta, n)`, check if `w` is above `bid_first(1, r, alpha, beta, n)` and if so return $10^{-6}$.

```{r}
# compute probability density for a winning bid from a first-price auction
w <- 0.5
compute_p_first_w(
  w = w, 
  r = r, 
  alpha = alpha, 
  beta = beta, 
  n = n
  )
upper <- 
  bid_first(
  x = 1, 
  r = r, 
  alpha = alpha, 
  beta = beta, 
  n = n
  )
compute_p_first_w(
  w = upper + 1, 
  r = r, 
  alpha = alpha, 
  beta = beta, 
  n = n
  )
```

7. Write a function `compute_loglikelihood_first_price_w(theta, df_first_w)` that computes the log-likelihood for a first-price auction winning bid data.

```{r}
# compute log-likelihood for winning bids for first-price auctions
compute_loglikelihood_first_price_w(
  theta = theta, 
  df_first_w = df_first_w
  )
```

8. Compare the value of the objective function around the true parameters.

```{r}
theta <- 
  c(
    alpha,
    beta
    )
# label
label <- 
  c(
    "\\alpha",
    "\\beta"
    )
label <- 
  paste(
    "$",
    label,
    "$", 
    sep = ""
    )
# compute the graph
graph <- 
  foreach (
    i = 1:length(theta)
    ) %do% {
  theta_i <- theta[i]
  theta_i_list <- 
    theta_i * seq(
      0.8,
      1.2,
      by = 0.05
      )
  objective_i <-
    foreach (
      j = 1:length(theta_i_list),
      .packages = c(
        "EmpiricalIO",
        "foreach",
        "magrittr"
        ),
      .combine = "rbind"
      ) %dopar% {
               theta_ij <- theta_i_list[j]
               theta_j <- theta
               theta_j[i] <- theta_ij
               objective_ij <- 
                 compute_loglikelihood_first_price_w(
                   theta_j,
                   df_first_w
                   )
               return(objective_ij)
             }
  df_graph <- 
    data.frame(
      x = as.numeric(theta_i_list), 
      y = as.numeric(objective_i)
      )
  g <-
    ggplot(
      data = df_graph,
      aes(
        x = x,
        y = y
        )
      ) +
    geom_point() +
    geom_vline(
      xintercept = theta_i, 
      linetype = "dotted"
      ) +
    ylab("objective function") + 
    xlab(TeX(label[i]))
  return(g)
}
saveRDS(
  graph,
  file = "data/a9/A9_first_parametric_graph.rds"
  )
```
```{r}
graph <- 
  readRDS(file = "data/a9/A9_first_parametric_graph.rds")
graph
```

9. Estimate the parameters by maximizing the log-likelihood. Set the lower bounds at zero. Use the Nelder-Mead method. Otherwise the parameter search can go to extreme values because of the discontinuity at the point where the upper limit is below the observed bid.

```{r}
result_first_parametric <-
  optim(
    par = theta,
    fn = compute_loglikelihood_first_price_w,
    df_first_w = df_first_w,
    method = "Nelder-Mead",
    control = list(fnscale = -1)
  )
saveRDS(
  result_first_parametric,
  file = "data/a9/A9_result_first_parametric.rds"
  )
```
```{r}
result_first_parametric <- 
  readRDS(file = "data/a9/A9_result_first_parametric.rds")
result_first_parametric
comparison <-
  data.frame(
    true = theta,
    estimate = result_first_parametric$par
  )
comparison
```

Finally, we non-parametrically estimate the distribution of the valuation using bid data from first-price auctions `df_first`.

10. Write a function `F_b(b)` that returns an empirical cumulative distribution at `b`. This can be obtained by using a function `ecdf`.
Also, write a function `f_b(b)` that returns an empirical probability density at `b`. This can be obtained by combining functions `approxfun` and `density`.

```{r}
# cumulative distribution
ggplot(
  df_first, 
  aes(x = b)
  ) + 
  stat_ecdf(color = "steelblue") +
  xlab("bid") + 
  ylab("cumulative distribution")
F_b <- ecdf(df_first$b)
F_b(0.4)
F_b(0.6)
# probability density
ggplot(
  df_first, 
  aes(x = b)
  ) + 
  geom_density(fill = "steelblue")
f_b <- approxfun(density(df_first$b))
f_b(0.4)
f_b(0.6)
```

The equilibrium distribution and density of the highest rival's bid are:
$$
H_b(b) := F_b(b)^{n - 1},
$$
$$
h_b(b) := (n - 1) f_b(b) F_b(b)^{n - 2}.
$$

11. Write a function `H_b(b, n, F_b)` and `h_b(b, n, F_b, f_b)` that return the equilibrium distribution and density of the highest rival's bid at point `b`.

```{r}
H_b(
  b = 0.4,
  n = n,
  F_b = F_b
  )
h_b(
  b = 0.4,
  n = n,
  F_b = F_b, 
  f_b
  )
```

When a bidder bids $b$, the implied valuation of her is:
$$
x = b + \frac{H_b(b)}{h_b(b)}.
$$

12. Write a function `compute_implied_valuation(b, n, r)` that returns the implied valuation given a bid. Let it return $x = 0$ if $b < r$, because we cannot know the value when the bid is below the reserve price.

```{r}
r <- df_first[1, "r"]
n <- df_first[1, "n"]
compute_implied_valuation(
  b = 0.4,
  n = n, 
  r = r, 
  F_b = F_b, 
  f_b = f_b
  )
```

13. Obtain the vector of implied valuations from the vector of bids and draw the empirical cumulative distribution. Overlay it with the true empirical cumulative distribution of the valuations.

```{r}
valuation_implied <- 
  df_first %>%
  dplyr::rowwise() %>%
  dplyr::mutate(
    x = compute_implied_valuation(
      b, 
      n,
      r,
      F_b, 
      f_b
      )
    ) %>%
  dplyr::ungroup() %>%
  dplyr::select(x) %>%
  dplyr::mutate(type = "estimate")
valuation_true <- 
  valuation %>%
  dplyr::select(x) %>%
  dplyr::mutate(type = "true")
valuation_plot <- 
  rbind(
    valuation_true,
    valuation_implied
    )
ggplot(
  valuation_plot,
  aes(
    x = x, 
    color = type
    )
  ) + 
  stat_ecdf()
```






